---
title: "BST203_lab1_2022"
date: '2022-07-25'
output: html_document
---

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(haven)
library(tidyr)
library(ggplot2)

# import the data set
df <- read_dta("lab1.dta")
```

### What is the overall mean logpasses? What is the sample size/ mean logpasses/variance in each group?
```{r}
# overall mean logpasses
mean(df$logpasses)

paste("The overall mean logpasses is ", round(mean(df$logpasses),3), ".", sep = "")
```
```{r}
# sample size/ mean logpasses/variance in each group

summ <- df %>% group_by(position) %>%
  summarize(sample_size = length(logpasses), 
            mean = mean(logpasses),
            variance = var(logpasses)) 

summ
```

### Are the assumptions met?

```{r}
ggplot() +
  geom_histogram(aes(x = logpasses),bins = 15, data = df) +
  facet_wrap(~position) +
  theme_bw()
```

Distributions appear normal

```{r}
ggplot() +
  geom_boxplot(aes(x = position, y = logpasses, fill = position), data = df) 

```

variance appear equal

### Estimate the within-group variance

```{r}
var_w <- summ %>% mutate( ns= (sample_size -1)*variance ) %>% ungroup() %>%
  mutate(var_w = sum(ns)/(sum(sample_size)-3))

var_w 

paste("Within-group variance is", round(unique(var_w$var_w),3))
```


### Estimate the between-group variance

```{r}
var_b <- summ %>% mutate(x_dff_n = (mean - mean(df$logpasses))^2*sample_size) %>%
  ungroup() %>%
  mutate(var_b = sum(x_dff_n)/2)

var_b

paste("Between-group variance is", round(unique(var_b$var_b),3))
```


### F-statistics

```{r}
paste("F-statistics is var_b/var_w =", round(unique(var_b$var_b/var_w$var_w),3))
```

### Find p-value for F-statistics

`df1` is the df for variance between groups (k-1)

`df2` is the df for variance within groups (n-k)

```{r}
pf(16.782, df1 = 2, df2 = dim(df)[1]-3, lower.tail = FALSE)
```


### anova function

```{r}
summary(aov(logpasses ~ position, df))
```








